---
title: "Sequences"
author: "Jasmine"
date: "12/02/2020"
output: html_document
---
# Known issues
currently has issues with 2 digit numbers (has the 0) but cannot edit on server unless on martone lab computer
```{r setup, include=FALSE}
library(tidyverse)
library(ptm)
library(googlesheets4)
```

# Sequences
Befor you start - to access files:
1. connect to ubc VPN
2. connect to server

#if we need to update the sequences run the `update_seq` function
```{r file access}
update_seq <- function(update = T){
  list.files("/Volumes/martonelab/Sequences/0_PTM/")
  
  list.files(p,pattern = "\\d{3}\\-\\d{3}", full.names = F,)
  
  photos <- list.files(p,pattern = "\\d{3}\\-\\d{3}", recursive = T)
  
  #dealing with sequences
  seq <- list.files("/Volumes/martonelab/Sequences/0_PTM",recursive = T)
  
  seq_clean <- as.data.frame(seq) %>%
    separate(seq, into = c("folder","file","fnm"),sep = "/", extra = "merge") %>%
    filter(!is.na(fnm)) %>%
    separate(fnm,
             into = c("ptm","gene","other"),
             sep = "_") %>%
    mutate(gene = str_to_lower(gene),
           gene = str_remove(gene,".ab"),
           gene = str_remove(gene,"1"))
  
  write_csv(seq_clean,paste0(Sys.Date(),"server_sequence.csv"))

}
```

# dealing with retrieved files

# open files
```{r}
seq_clean <- read_csv("output/2019-12-18 server_sequence.csv")
#!!! issues with being interperted as logicalin CO1 psba etc

ptm <- read_csv("./output/master_list_clean_photos.csv",
                col_types = "iccccccDcccccccccccccccccccnnccccc-iiccc")
  
#if just updating photos uncomment this line
#ptm <- masterlist()

 # prep file to be joined
ptm_s <- ptm %>%
  mutate(ptm = paste0("PTM", `PTM#`))
```

```{r echo=TRUE}
# clean up sequence names
genes <-  c("rbclrevnew", "gwsfn", "gwsrx", "psbar2", "psbaf", "F7553", "f|F57", "R1150K")

seq_rev <- seq_clean %>%
  filter(gene %in% genes) %>%
  distinct(gene, file, .keep_all = TRUE) 

# combine meta data to the files
piv <- left_join(ptm_s, seq_clean, by = "ptm") %>%
  mutate(
    gene = if_else(gene == "", "none", gene),
    gene = if_else(is.na(gene), "none", gene),
    present = 1) %>%
  pivot_wider(
    names_from = gene, values_from = present,
    values_fn = list(present = length)
  ) %>%
  distinct(`PTM#`, .keep_all = TRUE)


combine_FR <- function(forward, reverse){
  fr <- paste0(forward, reverse)
  str_remove_all(fr,"NA")
}

#putting in forward and reverse
piv_genetics <- piv %>% 
  mutate(psbar2 = if_else(psbar2 == 1, "R",""),
         psbaf = if_else(psbaf == 1, "F",""),
         p = combine_FR(psbaf,psbar2),
         gwsfn = if_else(gwsfn == 1, "F",""),
         gwsrx = if_else(gwsrx == 1, "R",""),
         g = combine_FR(gwsfn, gwsrx),
         rbclrevnew = if_else(rbclrevnew == 1, "R",""),
         f57 =  if_else(f57 == 1, "F",""),
         r = combine_FR(f57, rbclrevnew))

molecular_complete <- piv_genetics %>% 
  mutate(CO1 = if_else(is.na(CO1), g, CO1),
         psbA = if_else(is.na(psbA), p, psbA),
         rbcL = if_else(is.na(rbcL), r, rbcL))
```


# clean up and save
```{r}
last_col <- grep("Project", colnames(molecular_complete))

complete_table <- molecular_complete[1:last_col]

write_csv(complete_table,"./output/masterlist_clean_sequences.csv")
```

#working on it
- make sure sorting has not changed
- might need to make this a function 
- integrate the functional
```{r save data - sheets}
save_col <- cbind(complete_table$CO1, complete_table$psbA, complete_table$rbcL) %>% 
  as_tibble()
names(save_col) <- c("CO1 NEW","PSBA NEW", "RBCL NEW")

# CURRENTLY USING THE DUP for testing purposes
gs_ptm <- sheets_get("https://docs.google.com/spreadsheets/d/1vzVIT5gjQ0yCGwbAyqB4f8pltOkLfWz27TIFTxnm9hk/edit#gid=0")

sheets_edit(gs_ptm,
            save_col, 
            sheet = "All PTM data",
            range = "AL2") #last column
```